WIDS 2024 Final Project

### 2024-01-08 Code/R FIle.qmd
added more code comments, recoded party column in house data, moved filter for party from full_state_data to house clean script

```{r}
head(union_wages)
head(union_states)
head(house)
head(income_distribution)
head(minimum_wage)
```

@@

```{r}
summary(union_wages)
```

@@

```{r}
house <- house |>
  group_by(year, state, district) |> 
  filter(candidatevotes == max(candidatevotes)) |> 
  ungroup() |> 
  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
  select(!(state_fips:office)) |> #Remove unwanted columns
  select(!(stage:candidate)) |> 
  select(!(writein:fusion_ticket)) |> 
  rename(state_abbreviation = state_po)
```

@@

```{r}
union_states <- union_states |> 
  select(-state_census_code,
         -observations) |> 
  mutate(employment_count = employment * 1000, .keep = "unused") |> 
  mutate(union_employment_count = members * 1000, .keep = "unused") |> 
  mutate(collective_bargain_employment_count = covered * 1000, .keep = "unused") |> 
  rename(
    percent_union = p_members,
    percent_collective_bargain = p_covered
  )
```

```{r}
income_distribution <- income_distribution |> 
  select(-number) |> 
  select(!ends_with("moe")) |> 
  select(year:income_mean) |> 
  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> 
  rename(median_income = income_median,
         mean_income = income_mean)
```

```{r}
minimum_wage <- minimum_wage |> 
  select(Year:CPI.Average) |> 
  rename(year = Year,
         state = State,
         state_min_wage = State.Minimum.Wage,
         state_min_wage_2020_adj = State.Minimum.Wage.2020.Dollars,
         fed_min_wage = Federal.Minimum.Wage,
         fed_min_wage_2020_adj = Federal.Minimum.Wage.2020.Dollars,
         effective_min_wage = Effective.Minimum.Wage,
         effective_min_wage_2020_adj = Effective.Minimum.Wage.2020.Dollars,
         CPI_avg = CPI.Average
  )
```

```{r}
head(union_wages)
head(union_states)
head(house)
head(income_distribution)
head(minimum_wage)
```

```{r}
range(union_wages$year) #ntl
range(union_states$year) # state
range(house$year) # state
range(income_distribution$year) # ntl
range(minimum_wage$year) # state
```

@@

```{r}
full_state_data <- union_states |> 
  filter(sector == "Total") |> 
  select(-sector,
         -state) |> 
  left_join(house, by = c("year", "state_abbreviation")) |> 
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
  mutate(decade = paste0((year - year %% 10), "s"),
         party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
  left_join(minimum_wage, by = c("year", "state")) |> 
  group_by(year, state) |> 
  mutate(party_state_avaerage = mean(party_binary)) |>
  ungroup()
```

@@

>>>>

```{r}
# look at columns
head(union_wages)
head(union_states)
head(house)
head(income_distribution)
head(minimum_wage)
```

@@

```{r}
# look for NAs and skews
summary(union_wages)
```

@@

```{r}
house <- house |>
  group_by(year, state, district) |> # group for next line
  filter(candidatevotes == max(candidatevotes)) |> #only include candidates that won
  ungroup() |> 
  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
  select(!(state_fips:office)) |> #Remove unwanted columns
  select(!(stage:candidate)) |> 
  select(!(writein:fusion_ticket)) |> 
  rename(state_abbreviation = state_po) |>  # rename state abbreviation column
  mutate(party = case_when(
    str_detect(party, "DEMOCRAT") ~ "DEMOCRAT",
    str_detect(party, "REPUBLICAN") ~ "REUBLICAN",
    str_detect(party, "INDEPENDENT") ~ "INDEPENDENT"
  )) # Recode of party variable
```

@@

```{r}
union_states <- union_states |> 
  select(-state_census_code,
         -observations) |> # remove unwanted columns
  mutate(employment_count = employment * 1000, .keep = "unused") |> # column measured in 1000s, this fixes it
  mutate(union_employment_count = members * 1000, .keep = "unused") |> 
  mutate(collective_bargain_employment_count = covered * 1000,
         .keep = "unused") |> 
  rename(
    percent_union = p_members,
    percent_collective_bargain = p_covered
  ) #renamed columns
```

```{r}
income_distribution <- income_distribution |> 
  select(-number) |> # removing unwanted columns
  select(!ends_with("moe")) |> 
  select(year:income_mean) |> 
  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> # only include one line per year. needed bc of multiple sectors that are not being used in data analysis
  rename(median_income = income_median,
         mean_income = income_mean) # renaming columns
```

```{r}
minimum_wage <- minimum_wage |> 
  select(Year:CPI.Average) |> # selecting relevant columns
  rename(year = Year,
         state = State,
         state_min_wage = State.Minimum.Wage,
         state_min_wage_2020_adj = State.Minimum.Wage.2020.Dollars,
         fed_min_wage = Federal.Minimum.Wage,
         fed_min_wage_2020_adj = Federal.Minimum.Wage.2020.Dollars,
         effective_min_wage = Effective.Minimum.Wage,
         effective_min_wage_2020_adj = Effective.Minimum.Wage.2020.Dollars,
         CPI_avg = CPI.Average
  ) # renaming variables
```

```{r}
# seeing what we have left
head(union_wages)
head(union_states)
head(house)
head(income_distribution)
head(minimum_wage)
```

```{r}
# seeing what years we have
range(union_wages$year) #ntl
range(union_states$year) # state
range(house$year) # state
range(income_distribution$year) # ntl
range(minimum_wage$year) # state
```

@@

```{r}
full_state_data <- union_states |> 
  filter(sector == "Total") |> 
  select(-sector,
         -state) |> 
  left_join(house, by = c("year", "state_abbreviation")) |> 
  mutate(decade = paste0((year - year %% 10), "s"),
         party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
  left_join(minimum_wage, by = c("year", "state")) |> 
  group_by(year, state) |> 
  mutate(party_state_avaerage = mean(party_binary)) |>
  ungroup()
```

```{r}
union_states |> 
  filter(sector == "Total") |> 
  select(-sector,
         -state) |> 
  left_join(house, by = c("year", "state_abbreviation")) |> 
  count(year)
```

@@

```{r}
union_states |> 
  filter(sector == "Total") |> 
  select(-sector,
         -state) |> 
  left_join(house, by = c("year", "state_abbreviation")) |> 
  filter(party == "DEMOCRAT" | party == "REPUBLICAN" | party == "FOGLIETTA (DEMOCRAT)") |>
  count(year)
```
